Extended Data

### 2022-02-26 doc/manuscript/Dow_manuscript_Nature.Rmd
add correct table/figure names

Together, our results demonstrate that warmer spring temperatures in the temperate deciduous forests of eastern North America advance the timing of stem growth but have little effect on total annual growth (Figs. `r schematic_summary`- `r tree_rings`). 
The observed advance in stem growth under warmer springs parallels phenological advances observed for canopy foliage (Fig. `r DOY_timing`a-b)[@keenan_net_2014;@finzi_carbon_2020;@stuble_plant_2021], which makes sense given close coordination of budburst and xylogenesis[@delpierre_temperate_2016], and _NEE_[@keenan_net_2014;@finzi_carbon_2020]. 
**However, inconsistent with the concept that an earlier start to growth would increase annual woody productivity, we demonstrate that warmer springs either hasten the deceleration of stem expansion or otherwise failed to translate extended growing seasons into biologically significant increases in stem growth, and thereby have negligible effect on total annual growth for most species and locations (Fig. 3).**
We find statistically, but not biologically, significant relationships between total growth and spring temperature at both sites in our dendrometer band analysis. 
However, the slope of the relationships are in opposite directions, with warming causing decreasing stem growth at Harvard Forest but increasing stem growth at SCBI. 
We believe this difference is due to severe drought conditions in 1999 which obviously caused abnormal growth patterns in that year (pheno_tsens). 
Therefore, we suggest that our two sites represent two scenarios: 1. A forest which had no significant summer drought, but was only influenced by warming temperatures (SCBI). 2. A forest which experienced spring warming along with severe summer drought (Harvard Forest).

At SCBI, we see a very small increase in total stem growth (on average +0.05mm/$^\circ$C), suggesting that even in the absence of drought or incliment climate conditions during the growing season, stem growth is not enhanced in a meaningful way when spring temperatures are warmer, and peak stem growth is initiated earlier.  
At Harvard Forest, we see a more meaningful decrease in total growth in relation to warmer temperatures in the CTW (-0.15mm/$^\circ$C on average). We speculate that this result is likely a product of reductions in total stem growth related to drought damage in 1999. 
Due to the method of selection for CTW, years with significant departures from the norm can influence the analysis to a greater extent. Stem growth in 1999 was clearly impacted by climate conditions, and so the CTW selected by climwin maximizes the effect of these impacts (most likely the effects of drought) on stem growth. To examine this further, we reran our dendrometer band analysis using the month of May as the CTW. When we change the CTW to the month of May, we see results much closer to those found at SCBI (EDT...), with a similar advancement in growth timing but a much less pronounced effect of Tmax on total growth (-0.015mm/$^\circ$C on average, a tenfold decrease compared to the CTW optimized around the drought year). Interestingly, we once again see that there was no meaningful increase in total growth. 
The magnitude of the effect of the summer drought at Harvard Forest when compared to the relatively drought free period of SCBI illustrates the point that conditions during the peak growing season are much more influential to stem growth than conditions during the Spring. 
**We demonstrate that warmer springs either have little effect on, or even hasten, the deceleration of stem expansion and thereby have negligible effect on total annual stem growth for most species and locations (Fig. `r tree_rings`).** <!--repeated sentence/ CD made potential changes to the first one. See what you think. We should mention how 1999 was a drought year and also the warmest year in the CTW at HF for both wood types. This is potentially contributing to the negative correlations between CTW TMAX and total growth at HF. Maybe if there wasn't a severe drought in the warmest year we would see positive correlations like we do at SCBI. This is also probably worth mentioning when we discuss the influence of climate during peak growing season being greater than the spring climate. -->

>>>>

Together, our results demonstrate that warmer spring temperatures in the temperate deciduous forests of eastern North America advance the timing of stem growth but have little effect on total annual growth (Figs. `r schematic_summary`- `r tree_rings`). 
The observed advance in stem growth under warmer springs parallels phenological advances observed for canopy foliage (Fig. `r DOY_timing`a-b)[@keenan_net_2014;@finzi_carbon_2020;@stuble_plant_2021], which makes sense given close coordination of budburst and xylogenesis[@delpierre_temperate_2016], and _NEE_[@keenan_net_2014;@finzi_carbon_2020]. 
**However, inconsistent with the concept that an earlier start to growth would increase annual woody productivity, we demonstrate that warmer springs either hasten the deceleration of stem expansion or otherwise failed to translate extended growing seasons into biologically significant increases in stem growth, and thereby have negligible effect on total annual growth for most species and locations (Fig. 3).**
We find statistically, but not biologically, significant relationships between total growth and spring temperature at both sites in our dendrometer band analysis (`schematic_summary`). 
However, the slope of the relationships are in opposite directions, with warming causing decreasing stem growth at Harvard Forest but increasing stem growth at SCBI (`EDT_phenology_parameters`). 
We believe this difference is due to severe drought conditions in 1999 which obviously caused abnormal growth patterns in that year (`r EDF_pheno_Tmax_climwin`). 
Therefore, we suggest that our two sites represent two scenarios: 1. A forest which had no significant summer drought, but was only influenced by warming temperatures (SCBI). 2. A forest which experienced spring warming along with severe summer drought (Harvard Forest).

At SCBI, we see a very small increase in total stem growth (on average +0.05mm/$^\circ$C), suggesting that even in the absence of drought or inclement climate conditions during the growing season, spring warming does not enhance stem growth in a meaningful way.
At Harvard Forest, we see a decrease in total growth with warmer temperatures in the CTW (-0.15mm/$^\circ$C on average). We speculate that this result is likely a product of reductions in total stem growth related to drought effects in 1999. 
Due to the method of selection for CTW, years with significant departures from the norm can influence the analysis to a greater extent. 
Stem growth in 1999 was clearly impacted by climate conditions (`r EDF_pheno_Tmax_climwin`), and so the CTW selected by climwin maximizes the effect of these impacts (most likely the effects of drought) on stem growth. 
To examine this further, we reran our dendrometer band analysis using the month of May as the CTW. 
When we change the CTW to the month of May, we see results much closer to those found at SCBI, with a similar advancement in growth timing but a much less pronounced effect of $T_{max}$ on total growth (-0.015mm/$^\circ$C on average, a tenfold decrease compared to the CTW optimized around the drought year;`EDF_pheno_Tmax_AprilMay`). 
Once again see that there was no meaningful influence of spring temperature on total growth. 
In fact, the magnitude of the effect of the summer drought compared to the effects of spring $T_{max}$ illustrates the point that conditions during the peak growing season are much more influential to stem growth than conditions during the Spring. 
**We demonstrate that warmer springs either have little effect on, or even hasten, the deceleration of stem expansion and thereby have negligible effect on total annual stem growth for most species and locations (Fig. `r tree_rings`).** <!--repeated sentence/ CD made potential changes to the first one. See what you think. -->
